---
title: Credential Scraping
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Credential Scraping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(redcapcustodian)
```

## API Token Scraping and Fabricating

Using credential management tools requires `.env` files populated with database credentials for a each REDCap instance, the following examples use one for a production instance (`prod.env`) and one for a local instance (`local_1134.env`). Here is my `local_1134.env` file:

```bash
INSTANCE=local_1134
TIME_ZONE=America/New_York

# REDCap API endpoint, REDCap webroot + "/api/"
URI=https://localhost:11134/api/

# REDCap DB Credentials
REDCAP_DB_NAME=redcap
REDCAP_DB_HOST=127.0.0.1
REDCAP_DB_USER=redcap
REDCAP_DB_PASSWORD=redcap123
REDCAP_DB_PORT=3306
```

### Preparation

The following block creates a `credentials` directory and a sqlite database in which to store your credentials.

```{r, eval=FALSE}
library(redcapcustodian)
library(DBI)
library(tidyverse)
library(dotenv)

# fetching all extant API tokens and adding them to storage #################

dir.create("credentials")

# creates file if one does not exist
file_conn <- DBI::dbConnect(RSQLite::SQLite(), "credentials/credentials.db")

# SQLite friendly schema
credentials_sql <- "CREATE TABLE IF NOT EXISTS `credentials` (
  `redcap_uri` TEXT NOT NULL,
  `server_short_name` varchar(128) NOT NULL,
  `username` varchar(191) NOT NULL,
  `project_id` int(10) NOT NULL,
  `project_display_name` TEXT NOT NULL,
  `project_short_name` varchar(128) DEFAULT NULL,
  `token` varchar(64) NOT NULL,
  `comment` varchar(256) DEFAULT NULL
);
"

dbExecute(file_conn, credentials_sql)
```



### Scraping a server's API tokens and putting them in a local sqlite DB for use

Fetching all extant API tokens and adding them to storage 

```{r, eval = FALSE}
library(redcapcustodian)
library(DBI)
library(tidyverse)
library(dotenv)

# fetching all extant API tokens and adding them to storage #################

file_conn <- DBI::dbConnect(RSQLite::SQLite(), "credentials/credentials.db")

dbExecute(file_conn, credentials_sql)

load_dot_env("prod.env")

username <- "your_redcap_username"
source_conn <- connect_to_redcap_db()
source_credentials <- scrape_user_api_tokens(source_conn, username)

# alter credentials to match local schema
source_credentials_upload <- source_credentials %>%
  mutate(
    redcap_uri = Sys.getenv("URI"),
    server_short_name = tolower(Sys.getenv("INSTANCE"))
  ) %>%
  # remove duplicates
  anti_join(
    tbl(file_conn, "credentials") %>%
      collect()
  )

dbAppendTable(file_conn, "credentials", source_credentials_upload)
```

### Creating API tokens for all local projects

```{r, eval = FALSE}
library(redcapcustodian)
library(DBI)
library(tidyverse)
library(dotenv)

file_conn <- DBI::dbConnect(RSQLite::SQLite(), "credentials/credentials.db")

load_dot_env("local_1134.env")

# note, this will close any other connections
target_conn <- connect_to_redcap_db()

local_credentials <- scrape_user_api_tokens(target_conn, "admin")

# specify a subset of project_ids rather than making tokens for all
target_pids <- tbl(target_conn, "redcap_projects") %>%
  select(project_id) %>%
  filter(project_id > 15) %>%
  filter(!project_id %in% local(local_credentials$project_id)) %>%
  collect() %>%
  pull(project_id)

# create tokens individually
for(pid in target_pids) {
  set_project_api_token(target_conn, "admin", pid)
}

# gather newly created tokens and insert them into local storage
local_credentials <- scrape_user_api_tokens(target_conn, "admin")

local_credentials_upload <- local_credentials %>%
  mutate(
    redcap_uri = Sys.getenv("URI"),
    server_short_name = tolower(Sys.getenv("INSTANCE"))
  ) %>%
  # remove duplicates
  anti_join(
    tbl(file_conn, "credentials") %>%
      collect()
  )

dbAppendTable(file_conn, "credentials", local_credentials_upload)
```

## Syncing a production project to your local instance

After completing the previous 2 sections.

The following chink uses the credentials you have saved to copy project metadata from a source project ()

```{r, eval = FALSE}
library(redcapcustodian)
library(tidyverse)

file_conn <- DBI::dbConnect(RSQLite::SQLite(), "credentials/credentials.db")

source_username <- "your_production_username"

source_credentials <- tbl(file_conn, "credentials") %>%
  filter(source_username) %>%
  filter(server_short_name == "prod") %>%
  collect() %>%
  filter(
    str_detect(project_display_name, "The big important project")
  ) %>%
  unnest()

local_credentials <- tbl(file_conn, "credentials") %>%
  filter(username == "admin") %>%
  filter(server_short_name == "local_1134") %>%
  collect() %>%
  # adjust url to make REDCapR's validation processes happy
  mutate(redcap_uri = str_replace(redcap_uri, "https", "http")) %>%
  mutate(redcap_uri = str_replace(redcap_uri, "localhost", "127.0.0.1")) %>%
  filter(
    str_detect(project_display_name, "The big important project")
  ) %>%
  unnest()

sync_metadata(source_credentials, local_credentials, strip_action_tags = TRUE)
```

### Large data dictionaries and XDebug/REDCap Docker Compose

IF your REDCap instance is hosted on a machine (or docker container) using xdebug, the metadata may be too large, in which case you'll see an error message like this:
```
The REDCapR write/import metadata operation was not successful.  The error message was:\n\t\t\t\t<div class=\"red\" style=\"margin:20px 0px;max-width:700px;\">\n\t\t\t\t\t<b>REDCap crashed due to an unexpected PHP fatal error!</b><br><br>\n\t\t\t\t\t<b>Error message:</b> Uncaught Error: Xdebug has detected a possible infinite loop, and aborted your script with a stack depth of '256' frames in /var/www/html/redcap_v11.3.4/Classes/LogicParser.php:280\
```

Alter the following file `<your-rdc-instance>/redcap-overrides/web/php/80-xdebug.ini` and set an arbitrarily large stack frame limit, at least equal to the number of fields in your data dictionary 

```bash
# solution: https://stackoverflow.com/a/65997576/7418735
xdebug.max_nesting_level=<high-value>
```
