---
title: "Job Logging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Job Logging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval = T, include = F }
library(redcapcustodian)
library(tidyverse)
library(kableExtra)
dotenv::load_dot_env(here::here(".env"))
```

REDCap Custodian includes its own job logging facility. The system is designed to log each job run to a single row in a MYSQL database. With as little a four lines of code, you can log job success or failure along with a message. e.g.,

```{r minimal-example-of-success, eval = F, include = T, fig.cap = "A minimal example of logging a successful job"}
library(redcapcustodian)
library(dotenv)

init_etl("my_useful_script", log_db_drv = duckdb::duckdb())
# do some stuff ...

log_job_success("It worked!")
```

The code in above would produce a row in the logging database like that shown in below.

```{r mock_db_duckdb, eval = T, include = F}
init_etl("my_useful_script", log_db_drv = duckdb::duckdb())
log_job_success("It worked!")
```

```{r read-log-data, eval = T, include = F}
log_conn <- get_package_scope_var("log_con")

log_data <-
  dplyr::tbl(log_conn, "rcc_job_log") |>
  dplyr::filter(script_name == "my_useful_script") |>
  dplyr::collect()
```

```{r minimal-example-of-success-output, eval = T, echo = F, fig.cap = "Data written by `my_useful_script`"}
log_data |>
  kbl() |>
  kable_styling()
```

## Behind the scenes

That's a tiny script, but a lot is happening. Let's decode what's going on in those few lines. 

### `library(dotenv)`

REDCap Custodian uses the `dotenv` package to read files of name-value pairs into the environment of the script. By default it reads `.env` from the current directory. To make logging work correctly, that `.env` file will need a stanza like this:

``` ini
LOG_DB_NAME=rcc_log
LOG_DB_HOST=127.0.0.1
LOG_DB_USER=rcc_log
LOG_DB_PASSWORD=password
LOG_DB_PORT=9000
```

The "LOG_DB\_\*" strings are the environment variable names `redcapcustodian` uses to locate the MySQL DB where it stores the logging data. You need to use these names and the values need to point at a MySQL table on a MySQL server.

### `init_etl()`

`init_etl("my_useful_script")` sets the `script_name` env variable to "my_useful_script", sets the `script_run_time` env variable to the current date and time, and it verifies the connection to the database described by the "LOG_DB\_\*" environment variables. If the database is there, `init_etl()` quietly exits. If the database is not there you will see

``` r
Warning: Failed to connect to log DB: rcc_log_bad
The reason given was:
 Error: Failed to connect: Access denied for user 'rcc_log'@'%' to database 'rcc_log_bad'
```

### `log_job_success()`

`log_job_success("It worked!")` writes a row to the logging database like the example above.

## Failed jobs

When jobs fail you might want to log that condition, for that, call `log_job_failure()`. As with `log_job_success()`, you can provide a message to give more detail about what went wrong.

```r
log_job_failure("input dataframe was empty")
```

## More detailed messages

Logging "It worked!" or "input dataframe was empty" is fine, but often, you want to provide more detail about a job's activities. Chances are you have some useful facts in R objects already. It is a simple matter to assemble those objects into a list, convert the list object into a text string with using JSON libraries, and use that string as the job's message. If you wanted report some updates you made to the 'iris' database, you might log it like this:

```{r iris-updates, eval = T, echo = T}
init_etl("iris_updater")
# Do some stuff
# ...
iris_updates <- iris |> dplyr::sample_n(3) |> select(Sepal.Length, Species)
# ...
# success!
log_message <- dplyr::lst(iris_updates)
log_job_success(jsonlite::toJSON(log_message))
```

The `summary` value can contain any number of JSON-encoded objects. That said, plan them carefully. Ask what questions you will want to answer later after a job has run. The package authors have found the most common questions posed are "Why did this value change?", "Who (or what) set the value?", and "When did this change?" Store as little information as is needed to answer those questions succinctly. If you provide lots of different objects, it might make for a pile of log data to sift through later. We've found we can usually meet these needs by including minimal dataframes of the data written by the job. 

If you have multiple jobs writing to the same dataset, consider the value of self-consistency across those jobs. If every update operation of the _invoice_ table is logged in an object named _invoice_updates_, you can more easily search across those jobs.

## Reading back log data

Reading back the log data presents a few challenges. When you are answering the "Who? What? When? Why?" questions you often have to search across multiple log events to find the job run that made the change. You can filter the log data to the time frame or jobs of interest, but you will still have to search multiple job summaries to find the one that made the change you care about. 

If you encoded the job summary data with JSON, you'll need to decode it, but `jsonlite::fromJSON` is not vectorized. You'll have to iterate over the log records. Also, you'll want to unnest the job summary objects returned by `jsonlite::fromJSON`. REDCap Custodian provides the function `unnest_job_summary_data_json_object()` to help with these operations. e.g.,

```{r mock-updates-to-iris, eval = T, echo = F}
make_updates <- function() {
  Sys.sleep(1)
  iris_updates <- iris |> dplyr::sample_n(3) |> select(Sepal.Length, Species)
  log_job_success(jsonlite::toJSON(dplyr::lst(iris_updates)))  
}

make_updates()
make_updates()
```

```{r }
log_conn <- get_package_scope_var("log_con")

log_data <-
  dplyr::tbl(log_conn, "rcc_job_log") |>
  dplyr::filter(script_name == "iris_updater") |>
  dplyr::collect() |>
  dplyr::slice_tail(n = 3)
  
unnest_job_summary_data_json_object(log_data) |>
  unnest(cols = "iris_updates") |>
  kbl() |>
  kable_styling()
```

```{r, eval = T, include = F}
DBI::dbDisconnect(log_conn)
```
